### 2018-07-29 doc/Paper/authors.tex
migration to desc

% Author list file generated with: mkauthlist 1.2.3 
% mkauthlist -j mnras -f -c LSST Dark Energy Science Collaboration --cntrb contributions.tex authors.csv authors.tex

\author[LSST Dark Energy Science Collaboration]{
\parbox{\textwidth}{
\Large
David Alonso$^{1,2}$,
Javier Sanchez$^{3}$,
An\v{z}e Slosar$^{4}$
\begin{center} (LSST Dark Energy Science Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{
%\scriptsize
$^{1}$ School of Physics and Astronomy, Cardiff University, The Parade, Cardiff, CF24 3AA, United Kingdom\\
$^{2}$ Department of Physics, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, United Kingdom\\
$^{3}$ Department of Physics and Astronomy, University of California, Irvine, CA 92697, USA\\
$^{4}$ Brookhaven National Laboratory, Physics Department, Upton, NY 11973, USA\\
}
}